### 2019-04-27 Thesis/promonads.tex
Monthly commit.

\documentclass{article}

@@

\section{Promonads and parametric promonads}
\label{SecPromonads}

The purpose of this section is to shine some light on the definition of the \Mellies category, showing why it is natural to think about $[\X,\Set]$-enriched categories and why the \Mellies category may be considered analogous to the Kleisli category.

@@

The first half of this section is, in the interests of completeness, fairly technical, and may be skimmed over on a first reading.  
In section \ref{SecEndoprofunctors}, we introduce the multicategory of endoprofunctors on a category $\C$, which generalizes the monoidal category of endofunctors on $\C$.  
As monoids in $\End(\C)$ are called monads on $\C$, so we will call monoids in $\Endoprof(\C)$ \emph{promonads} on $\C$.  
We will observe that a promonad may be regarded as a sort of category, and that the Kleisli category may be characterized as the embedding of monads on $\C$ into promonads on $\C$.

@@

\subsection{Multicategories}

@@

\subsection{Representable multicategories}

@@

\subsection{Product and unit multicategories}

@@

\subsection{Multifunctors and multinatural transformations}

@@

\subsection{Monoids in multicategories}

@@

\subsection{Categories enriched over multicategories}

@@

\subsection{Multicategory-enriched functors and natural transformations}

@@

\subsection{The categories enriched over a symmetric multicategory form a multicategory}

@@

\subsection{Change of base}

@@

\subsection{Closed multicategories}

@@

\subsection{The multicategory of endoprofunctors}
\label{SecEndoprofunctors}

@@

\begin{SidewaysFigure}
  \[
    \begin{tikzcd}[ampersand replacement=\&, row sep=80pt, column sep=30pt]
      %
        \& F_1,\cdots,F_n \arrow[dr, "{\phi^{(1)},\cdots,\phi^{(m)}}"]
          \&
            \& \\
      %
        \& F_1,\cdots,F_{k_1},\C,\cdots,\C,F_{k_{m-1}},\cdots,F_n \arrow[u, "{\id_{F_1},\cdots,\id_{F_{k_1}},p,\cdots,p,\id_{F_{k_{m-1}+1}},\cdots,\id_{F_n}}"' {description, near start, xshift=6ex}] \arrow[dr, "{\phi^{(1)},\id,\cdots,\id,\phi^{(n)}}" description]
          \& G_1,\cdots,G_m \arrow[dr, "\psi"]
            \& \\
      F_1,\C_1,F_2,\cdots,F_{n-1},\C,F_n \arrow[uur, "{\id,p,\cdots,p}"] \arrow[ddr, "{p,\cdots,p,\id}"'] \arrow[ur, "{\id,p,\cdots,p,\id_{\C},\cdots,\id_{\C},\id,\cdots,p}"' {description, xshift=6ex}] \arrow[dr, "{p,\cdots,p,\id,\id_{\C},\cdots,\id_{\C},p,\cdots,p,\id}" {description, xshift=6ex}]
        \&
          \& G_1,\C,G_2,\cdots,G_{m-1},\C,G_m \arrow[u, "{\id,p,\cdots,p}" description] \arrow[d, "{p,\cdots,p,\id}" description]
            \& H \\
      %
        \& F_1,\cdots,F_{k_1},\C,\cdots,\C,F_{k_{m-1}},\cdots,F_n \arrow[d, "{\id_{F_1},\cdots,\id_{F_{k_1}},p,\cdots,p,\id_{F_{k_{m-1}+1}},\cdots,\id_{F_n}}"' {description, near start, xshift=6ex}] \arrow[ur, "{\phi^{(1)},\id,\cdots,\id,\phi^{(n)}}" description]
          \& G_1,\cdots,G_m \arrow[ur, "\psi"']
            \& \\
      %
        \& F_1,\cdots,F_n \arrow[ur, "{\phi^{(1)},\cdots,\phi^{(m)}}"']
          \&
            \& \\
    \end{tikzcd}
    \]
  \caption{Proof that extranaturality is preserved by composition.  
  Commutativity of the central square is by extranaturality of the $\phi^{(i)}$, while that of the four-cornered triangle at the right is by extranaturality of $\psi$.  
  The triangles on the left commute automatically, while the parallelograms at the top and the bottom commute by naturality of the $\phi^{(i)}$.}
  \label{FigExtranaturalityComposition}
\end{SidewaysFigure}

@@

\subsection{Functors are a special case of profunctors}

@@

\subsection{Promonads are categories}

@@

\subsection{The multicategory of functors}

@@

\subsection{Monoids on functors are multifunctors}
It might seem strange that the objects of the multicategory of functors are ordinary functors, ignoring the monoidal structure of $\X$ and the multicategory structure of $\M$.  
One way to make sense of this fact is to note that an object of a category $\C$ is the same thing as a functor
\[
  1 \to \C\,.
  \]
In the same way, perhaps the correct way to think of an `element' of a multicategory is that it is a \emph{multifunctor}
\[
  1 \to \M\,;
  \]
i.e., a monoid in $\M$.

@@

\subsection{Two perspectives on monoids in $\Set$}

We now come to our main result of the section.  
By way of motivation, consider the two ways of generalizing the notion of an internal monoid in $\Set$.

>>>>

\documentclass{report}[11pt]

@@

\chapter{Promonads and parametric promonads}
\label{SecPromonads}

The purpose of this chapter is to shine some light on the definition of the \Mellies category, showing why it is natural to think about $[\X,\Set]$-enriched categories and why the \Mellies category may be considered analogous to the Kleisli category.

@@

The first half of this chapter is, in the interests of completeness, fairly technical, and may be skimmed over on a first reading.  
In chapter \ref{SecEndoprofunctors}, we introduce the multicategory of endoprofunctors on a category $\C$, which generalizes the monoidal category of endofunctors on $\C$.  
As monoids in $\End(\C)$ are called monads on $\C$, so we will call monoids in $\Endoprof(\C)$ \emph{promonads} on $\C$.  
We will observe that a promonad may be regarded as a sort of category, and that the Kleisli category may be characterized as the embedding of monads on $\C$ into promonads on $\C$.

@@

\section{Multicategories}

@@

\section{Representable multicategories}

@@

\section{Product and unit multicategories}

@@

\section{Multifunctors and multinatural transformations}

@@

\section{Monoids in multicategories}

@@

\section{Categories enriched over multicategories}

@@

\section{Multicategory-enriched functors and natural transformations}

@@

\section{The categories enriched over a symmetric multicategory form a multicategory}

@@

\section{Change of base}

@@

\section{Closed multicategories}

@@

\section{The multicategory of endoprofunctors}
\label{SecEndoprofunctors}

@@

\begin{SidewaysFigure}
  \[
    \begin{tikzcd}[ampersand replacement=\&, row sep=80pt, column sep=30pt]
      %
        \& F_1,\cdots,F_n \arrow[dr, "{\phi^{(1)},\cdots,\phi^{(m)}}"]
          \&
            \& \\
      %
        \& F_1,\cdots,F_{k_1},\C,\cdots,\C,F_{k_{m-1}},\cdots,F_n \arrow[u, "{\id_{F_1},\cdots,\id_{F_{k_1}},p,\cdots,p,\id_{F_{k_{m-1}+1}},\cdots,\id_{F_n}}"' {description, near start, xshift=6ex}] \arrow[dr, "{\phi^{(1)},\id,\cdots,\id,\phi^{(n)}}" description]
          \& G_1,\cdots,G_m \arrow[dr, "\psi"]
            \& \\
      F_1,\C_1,F_2,\cdots,F_{n-1},\C,F_n \arrow[uur, "{\id,p,\cdots,p}"] \arrow[ddr, "{p,\cdots,p,\id}"'] \arrow[ur, "{\id,p,\cdots,p,\id_{\C},\cdots,\id_{\C},\id,\cdots,p}"' {description, xshift=6ex}] \arrow[dr, "{p,\cdots,p,\id,\id_{\C},\cdots,\id_{\C},p,\cdots,p,\id}" {description, xshift=6ex}]
        \&
          \& G_1,\C,G_2,\cdots,G_{m-1},\C,G_m \arrow[u, "{\id,p,\cdots,p}" description] \arrow[d, "{p,\cdots,p,\id}" description]
            \& H \\
      %
        \& F_1,\cdots,F_{k_1},\C,\cdots,\C,F_{k_{m-1}},\cdots,F_n \arrow[d, "{\id_{F_1},\cdots,\id_{F_{k_1}},p,\cdots,p,\id_{F_{k_{m-1}+1}},\cdots,\id_{F_n}}"' {description, near start, xshift=6ex}] \arrow[ur, "{\phi^{(1)},\id,\cdots,\id,\phi^{(n)}}" description]
          \& G_1,\cdots,G_m \arrow[ur, "\psi"']
            \& \\
      %
        \& F_1,\cdots,F_n \arrow[ur, "{\phi^{(1)},\cdots,\phi^{(m)}}"']
          \&
            \&
    \end{tikzcd}
    \]
  \caption{Proof that extranaturality is preserved by composition.  
  Commutativity of the central square is by extranaturality of the $\phi^{(i)}$, while that of the four-cornered triangle at the right is by extranaturality of $\psi$.  
  The triangles on the left commute automatically, while the parallelograms at the top and the bottom commute by naturality of the $\phi^{(i)}$.}
  \label{FigExtranaturalityComposition}
\end{SidewaysFigure}

@@

\section{Functors are a special case of profunctors}

@@

\section{Promonads are categories}

@@

\section{The multicategory of functors}

@@

\section{Monoids on functors are multifunctors}
It might seem strange that the objects of the multicategory of functors are ordinary functors, ignoring the monoidal structure of $\X$ and the multicategory structure of $\M$.  
One way to make sense of this fact is to note that an object of a category $\C$ is the same thing as a functor
\[
  1 \to \C\,.
  \]
In the same way, perhaps the correct way to think of an `element' of a multicategory is that it is a \emph{multifunctor}
\[
  1 \to \M\,;
  \]
i.e., a monoid in $\M$.

@@

\section{Two perspectives on monoids in $\Set$}

We now come to our main result of the chapter.  
By way of motivation, consider the two ways of generalizing the notion of an internal monoid in $\Set$.

### 2019-04-27 Thesis/thesis.tex
Monthly commit.

In addition, we will specialize to reader monads.  
The rationale behind this is twofold: firstly, we do not wish to assume too much about the underlying category $\G$ beyond the fact that it gives us a good model of Idealized Algol.  
Secondly, in order to get a compositional semantics for the $\lambda$-calculus, it is important that the Kleisli categories we consider are themselves Cartesian closed.

@@

\begin{definition}
  Fix some constant value $\top \in X$ (the precise value does not matter).
  We inductively define terms in context $\tr_u\from \nat \to X$ of \emph{ordinary deterministic} Idealized Algol for each $u\in X^*$ as follows.
  \begin{mathpar}
    \tr_\epsilon = \lambda n.\top
    \and
    \tr_{xu} = \lambda n.\neww (\lambda v.v\gets n;\IfO !v \Then x \Else \tr_u (\pred !v))
  \end{mathpar}
\end{definition}

@@

@@

First recall the statement of Computational Adequacy for $\G$:

\begin{proposition}
  Let $M\from \com$ be a closed term of Idealized Algol and suppose that
  \[
    ,()\ts M\converges \skipp,()\,.
    \]
  Then $\deno{M} \ne \bot$.
\end{proposition}

@@

\begin{proposition}[Soundness]
  Let $M\from\com$ be a closed term of \IAX, let $u\in X^*$ be a sequence and suppose that
  \[
    ,()\ts M\converges_u\skipp,()\,.
    \]
  Let the denotation $\deno{M} \from 1 \to \com$ in $\G_X$ be considered as a morphism $1 \to (X \to \bC)$ in $\G$.
  Then the composite
  \[
    1 \xrightarrow{\deno{M}}
    (X \to \bC) \xrightarrow{\eta_u}
    (\Var \to \bN) \xrightarrow{\deno{\neww}}
    \bN \xrightarrow{t_{|u|}}
    \bC
    \]
  is not equal to $\bot$.
  \label{PropKleisliSoundness}
\end{proposition}
\begin{proof}
  Since the $\beta$ rule is valid in $\G_X$, this composite is equal to the denotation of the term
  \[
    \test_{|u|}(\new (\lambda v.M[v\gets \suc !v;\tr_{u^\top} !v/\ask_X];!v))
    \]
  in IA.
  By the adequacy result for Idealized Algol, it suffices to show that this term converges to $\skipp$; i.e., that the term
  \[
    \new (\lambda v.M[v\gets\suc !v;\tr_{u^\top} !v/\ask_X];!v)
    \]
  converges to $|u|$ in IA.

  We can prove this using the following derivation tree.
  \[
    \inferrule*
    {
      \inferrule*[sep=1.7em]
      {
        \inferrule*[left={Lem. \ref{LemSoundness}}]
        {
        }
        {
          v,(v\mapsto 0)\ts M[v\gets\suc!v;\tr_{u^\top} !v/\ask_x]\converges \skipp,(v\mapsto |u|)
        }
        \\
        \inferrule*
        {
          \inferrule*
          {
          }
          {
            v,(v\mapsto |u|) \ts v \converges v,(v\mapsto |u|)
          }
        }
        {
          v,(v\mapsto |u|)\ts !v \converges |u|,(v\mapsto |u|)
        }
      }
      {
        v,(v\mapsto 0) \ts M[v\gets \suc !v;\tr_{u^\top} !v/\ask_X];v \converges |u|,(v\mapsto |u|)
      }
    }
    {
      ,()\ts\new (\lambda v.M[v\gets \suc !v;\tr_{u^\top} !v/\ask_X];!v) \converges |u|,()
    }\qedhere
    \]
\end{proof}

@@

If $\G$ is the category of games and visible strategies, then the statements of Proposition \ref{PropKleisliSoundness} (and our later Adequacy and Full Abstraction results) become clearer.  
Observe that if $\sigma \from 1 \to (X \to \bC)$ is a strategy in $\G$ (considered as a strategy for $\oc X \implies \bC$, then the maximal plays in the interaction
\[
  \sigma || (\eta_u;\deno{\neww})
  \]
take the form
\[
  \begin{array}{ccc}
    X         & \bC & \bN \\
              &     &  q  \\
              &  q  &     \\
    q         &     &     \\
    u^{(0)}   &     &     \\
    \vdots    &     &     \\
    q         &     &     \\
    u^{(k-1)} &     &     \\
              &  a  &     \\
              &     &  k  \\
  \end{array}\,,
  \]
for $k\le |u|$ or
\[
  \begin{array}{ccc}
    X         & \bC & \bN \\
              &     &  q  \\
              &  q  &     \\
    q         &     &     \\
    u^{(0)}   &     &     \\
    \vdots    &     &     \\
    q         &     &     \\
    u^{(|u|-1)}&     &     \\
    q         &     &     \\
    \top      &     &     \\
    \vdots    &     &     \\
    q         &     &     \\
    \top      &     &     \\
              &  a  &     \\
              &     &  k  \\
  \end{array}\,,
  \]
for $k>|u|$, where the component in $X,\bC$ is a valid play of $\sigma$.
Moreover, the strategy $t_n$ is the one with maximal plays of the form
\[
  \begin{array}{cc}
    \bN & \bC \\
        &  q  \\
    q   &     \\
    n   &     \\
        &  a
  \end{array}
  \]
or
\[
  \begin{array}{cc}
    \bN & \bC \\
        &  q  \\
    q   &     \\
    m   &
  \end{array}
  \]
for $m\ne n$.

@@

We can now prove computational adequacy for our model.

\begin{proposition}[Computational adequacy]
  Let $M\from \com$ be a closed term of \IAX.  
  Consider the denotation $\deno M\from 1 \to \bC$ in $\G_X$ as a morphism $1 \to (X \to \bC)$ in $\G$.  
  Let $u\in X^*$ be a sequence and suppose that the composite
  \[
    1 \xrightarrow{\deno{M}}
    (X \to \bC) \xrightarrow{\eta_u}
    (\Var \to \bN) \xrightarrow{\deno{\neww}}
    \bN \xrightarrow{t_{|u|}}
    \bC
    \]
  is not equal to $\bot$.  
  Then
  \[
    ,()\ts M\converges_u \skipp,()\,.
    \]
  \label{PropKleisliAdequacy}
\end{proposition}
\begin{proof}
  As before, the composite given in the statement is the denotation of the term
  \[
    \test_{|u|} (\new (\lambda v.M[v\gets \suc !v;\tr_{u^\top} !v/\ask_X];!v))\\,.
    \]
  By the adequacy result for Idealized Algol, the fact that this denotation is not equal to $\bot$ means that the term converges to $\skipp$, from which we can deduce that
  \[
    \new (\lambda v.M[v\gets \suc !v;\tr_{u^\top} !v/\ask_X];!v
    \]
  converges to $|u|$.

  It is easy to see that this is equivalent to derivability of the following relation in Idealized Algol.
  \[
    v,(v\mapsto 0) \ts M[v\gets \suc !v;\tr_{u^\top} !v/\ask_X] \converges \skipp,(v\mapsto |u|)
    \]
  Now $u$ is the subsequence of $u^\top$ going from position $1$ to position $|u|$.  
  So Lemma \ref{LemAdequacy} tells us that we must have
  \[
    ,()\ts M \converges_u \skipp,()
    \]
  in \IAX.
\end{proof}

@@

\begin{theorem}[Full abstraction]
  Let $M,N\from T$ be closed terms of \IAX.  
  Then $M,N$ are observationally equivalent -- i.e., for all contexts $C[-]\from\com$ of \IAX with a hole of type $T$ and for all sequences $u\in X^*$, 
  \[
    ,()\ts C[M] \converges_u \skipp,() \Longleftrightarrow ,()\ts C[N] \converges_u \skipp,()\,\text{--}
    \]
  if and only if $\deno{M}\sim\deno{N}$.
  \label{TheKleisliFullAbstraction}
\end{theorem}
\begin{proof}
  First, suppose that $\deno{M}\sim\deno{N}$.  
  Let $C[-]\from\com$ be a context with a hole of type $T$.  
  Then the denotation of $t\ts C[t]$ is a morphism $\alpha\from \deno{T} \to \bC$.
  Moreover, the denotation of $C[M]$ is the composite $\deno{M};\alpha$ and that of $C[N]$ is the composite $\deno{N};\alpha$ by functional completeness.

@@

\begin{corollary}
  Let $M\from\com$ be a closed term of \IAX.  
  Consider the denotation $\deno M\from 1 \to \bC$ in $\G_X$ as a morphism $1 \to (X \to \bC)$ in $\G$.  
  Then there exists some sequence $u\in X^*$ such that the composite
  \[
    1 \xrightarrow{\deno M}
    (X \to \bC) \xrightarrow{\eta_u}
    (\Var \to \bN) \xrightarrow{\deno{\neww}}
    \bN \xrightarrow{t_{|u|}}
    \bC
    \]
  is not equal to $\bot$, if and only if
  \[
    ,()\ts M \converges \skipp,()\,.
    \]
  \label{CorMayAdequacy}
\end{corollary}

@@

\begin{theorem}
  Let $M,N\from T$ be closed terms of \IAX. Then $M,N$ are may-observationally equivalent -- i.e., for all contexts $C[-]\from \com$ of \IAX with a hole of type $T$, \[
    ,()\ts C[M] \converges \skipp,() \Leftrightarrow ,()\ts C[N] \converges \skipp,()\,\text{--}
    \]
  if and only if $\deno M \sim_{\may} \deno N$.
\end{theorem}
\begin{proof}
  Suppose that $\deno M\sim_{\may}\deno N$.  
  Let $C[-]$ be a context of \IAX, interpreted as a morphism $\alpha \from \deno T \to \bC$.

  Suppose that $,()\ts C[M] \converges \skipp,()$.  
  So there is some sequence $u\in X^*$ such that $,()\ts C[M] \converges_u \skipp,()$ and therefore the composite
  \[
    1 \xrightarrow{\deno M;\alpha}
    (X \to \bC) \xrightarrow{\eta_u}
    (\Var \to \bN) \xrightarrow{\deno{\neww}}
    \bN \xrightarrow{t_{|u|}}
    \bC
    \]
  is not equal to $\bot$.  
  Therefore, there exists some $v\in X^*$ such that the composite
  \[
    1 \xrightarrow{\deno N;\alpha}
    (X \to \bC) \xrightarrow{\eta_v}
    (\Var \to \bN) \xrightarrow{\deno{\neww}}
    \bN \xrightarrow{t_{|v|}}
    \bC
    \]
  Therefore, $,()\ts C[N] \converges_v \skipp,()$ and so $,()\ts C[N] \converges \skipp,()$.
  The reverse direction is identical.

@@

\begin{example}
  $,()\ts \Y(\lambda x.x)\diverges$ because it has the following infinite derivation tree.
  \[
    \inferrule*
    {
      \inferrule*
      {
        \inferrule*
        {
        }
        {
          ,()\ts (\lambda x.x) \converges (\lambda x.x),()
        }
        \\
        \inferrule*
        {
          \vdots
        }
        {
          ,() \ts \Y(\lambda x.x)\diverges
        }
      }
      {
        ,()\ts (\lambda x.x)(\Y(\lambda x.x)) \diverges
      }
    }
    {
      ,()\ts \Y(\lambda x.x)\diverges
    }
    \]
\end{example}

@@

\begin{corollary}
  Let $M\from \com$ be a closed term of \IAX and consider the denotation $\deno M \from 1 \to \bC$ in $\G_X$ as a morphism $1 \to (X \to \bC)$ in $\G$.
  Then $,()\ts M\mustconverge$ if and only if $\deno{M}\downarrow_{\must}$.
  \label{CorAdeqMust}
\end{corollary}

@@

We have shown the following adequacy result.
\begin{corollary}
  Let $M\from \com$ be a closed term of \IAX and consider the denotation $\deno M \from 1 \to \bC$ in $\G_X$ as a morphism $X \to \bC$ in $\G$.
  Then $,()\ts M\mustconverge$ if and only if this strategy is winning.
\end{corollary}

>>>>

In addition, we will specialize to reader monads.  
The rationale behind this is twofold: firstly, we do not wish to assume too much about the underlying category $\G$ beyond the fact that it gives us a good model of Idealized Algol.  
Reader monads can be constructed for any object inside any Cartesian closed category.
Secondly, in order to get a compositional semantics for the $\lambda$-calculus, it is important that the Kleisli categories we consider are themselves Cartesian closed.  
Theorem \ref{FunctionalCompletenessCcc} tells us that this is the case for the Kleisli categories of reader monads on categories that are themselves Cartesian closed, but this may not be the case for other monads we can build from the Cartesian closed structure on our base category, such as state monads.

@@

\begin{definition}
  We inductively define terms in context $\tr_u\from \nat \to X$ of \emph{ordinary deterministic} Idealized Algol for each $u\in X^*$ as follows.
  \begin{mathpar}
    \tr_\epsilon = \lambda n.\Omega
    \and
    \tr_{xu} = \lambda n.\neww (\lambda v.v\gets n;\IfO !v \Then x \Else \tr_u (\pred !v))
  \end{mathpar}
\end{definition}

@@

            \inferrule*[sep=5pt]

@@

First recall the statement of Computational Adequacy for $\G$ (Theorem \ref{TheComputationalAdequacyIA} if $\G$ is the category of games):

\begin{theorem}
  Let $M\from \com$ be a closed term of Idealized Algol.  
  Then
  \[
    \blank,()\ts M\converges \skipp,()\,.
    \]
  if and only if $\deno{M} \ne \bot$.
\end{theorem}

@@

\begin{definition}
  Let $\sigma \from 1 \to \bC$ be a morphism in $\G_X$, considered as a morphism $(X \to \bC)$ in $\G$.  
  We say that $\sigma\downarrow_u$ if the composite
  \[
    1 \xrightarrow{\sigma}
    (X \to \bC) \xrightarrow{\eta_u}
    (\Var \to \bN) \xrightarrow{\deno{\neww}}
    \bN \xrightarrow{t_{|u|}}
    \bC
    \]
  is not equal to $\bot$.
\end{definition}

\begin{proposition}
  Let $M\from\com$ be a closed term of \IAX, let $u\in X^*$ be a sequence and suppose that
  \[
    \blank,()\ts M\converges_u\skipp,()\,.
    \]
  Let the denotation $\deno{M} \from 1 \to \com$ in $\G_X$ be considered as a morphism $1 \to (X \to \bC)$ in $\G$.
  Then $\deno{M}\downarrow_u$; i.e., the composite
  \[
    1 \xrightarrow{\deno{M}}
    (X \to \bC) \xrightarrow{\eta_u}
    (\Var \to \bN) \xrightarrow{\deno{\neww}}
    \bN \xrightarrow{t_{|u|}}
    \bC
    \]
  is not equal to $\bot$.
  \label{PropKleisliSoundness}
\end{proposition}
\begin{proof}
  Since the $\beta$ rule is valid in $\G_X$, this composite is equal to the denotation of the term
  \[
    \test_{|u|}(\new (\lambda v.M[v\gets \suc !v;\tr_{u^\top} !v/\ask_X];!v))
    \]
  in IA.
  By the adequacy result for Idealized Algol, it suffices to show that this term converges to $\skipp$; i.e., that the term
  \[
    \new (\lambda v.M[v\gets\suc !v;\tr_{u^\top} !v/\ask_X];!v)
    \]
  converges to $|u|$ in IA.

  We can prove this using the following derivation tree.
  \tiny
  \[
    \inferrule*
    {
      \inferrule*[sep=0em]
      {
        \inferrule*[left={Lem. \ref{LemSoundness}}]
        {
        }
        {
          v,(v\mapsto 0)\ts M[v\gets\suc!v;\tr_{u^\top} !v/\ask_x]\converges \skipp,(v\mapsto |u|)
        }
        \\
        \inferrule*
        {
          \inferrule*
          {
          }
          {
            v,(v\mapsto |u|) \ts v \converges v,(v\mapsto |u|)
          }
        }
        {
          v,(v\mapsto |u|)\ts !v \converges |u|,(v\mapsto |u|)
        }
      }
      {
        v,(v\mapsto 0) \ts M[v\gets \suc !v;\tr_{u^\top} !v/\ask_X];v \converges |u|,(v\mapsto |u|)
      }
    }
    {
      \blank,()\ts\new (\lambda v.M[v\gets \suc !v;\tr_{u^\top} !v/\ask_X];!v) \converges |u|,()
    }\qedhere
    \]
  \normalsize
\end{proof}

@@

If $\G$ is the category of games and visible strategies, then the statements of Proposition \ref{PropKleisliSoundness} (and our later Adequacy and Full Abstraction results) become clearer.  
Observe that if $\sigma \from 1 \to (X \to \bC)$ is a strategy in $\G$ (considered as a strategy for $\oc X \implies \bC$, then the maximal plays in the interaction
\[
  \sigma || (\eta_u;\deno{\neww})
  \]
take the form
\[
  \begin{array}{ccc}
    X         & \bC & \bN \\
              &     &  q  \\
              &  q  &     \\
    q         &     &     \\
    u^{(0)}   &     &     \\
    \vdots    &     &     \\
    q         &     &     \\
    u^{(k-1)} &     &     \\
              &  a  &     \\
              &     &  k  \\
  \end{array}\,,
  \]
for $k\le |u|$, where the component in $X,\bC$ is a valid play of $\sigma$.
Moreover, the strategy $t_n$ is the one with maximal plays of the form
\[
  \begin{array}{cc}
    \bN & \bC \\
        &  q  \\
    q   &     \\
    n   &     \\
        &  a
  \end{array}
  \]
or
\[
  \begin{array}{cc}
    \bN & \bC \\
        &  q  \\
    q   &     \\
    m   &
  \end{array}
  \]
for $m\ne n$.

@@

We can now prove computational adequacy for our model.
We have proved one direction already in Proposition \ref{PropKleisliSoundness}, so it suffices to prove the other direction.

\begin{proposition}[Computational adequacy]
  Let $M\from \com$ be a closed term of \IAX.  
  Consider the denotation $\deno M\from 1 \to \bC$ in $\G_X$ as a morphism $1 \to (X \to \bC)$ in $\G$.  
  Let $u\in X^*$ be a sequence and suppose that the composite
  \[
    1 \xrightarrow{\deno{M}}
    (X \to \bC) \xrightarrow{\eta_u}
    (\Var \to \bN) \xrightarrow{\deno{\neww}}
    \bN \xrightarrow{t_{|u|}}
    \bC
    \]
  is not equal to $\bot$.  
  Then
  \[
    \blank,()\ts M\converges_u \skipp,()\,.
    \]
  \label{PropKleisliAdequacy}
\end{proposition}
\begin{proof}
  As before, the composite given in the statement is the denotation of the term
  \[
    \test_{|u|} (\new (\lambda v.M[v\gets \suc !v;\tr_{u^\top} !v/\ask_X];!v))\\,.
    \]
  By the adequacy result for Idealized Algol, the fact that this denotation is not equal to $\bot$ means that the term converges to $\skipp$, from which we can deduce that
  \[
    \new (\lambda v.M[v\gets \suc !v;\tr_{u^\top} !v/\ask_X];!v
    \]
  converges to $|u|$.

  It is easy to see that this is equivalent to derivability of the following relation in Idealized Algol.
  \[
    v,(v\mapsto 0) \ts M[v\gets \suc !v;\tr_{u^\top} !v/\ask_X] \converges \skipp,(v\mapsto |u|)
    \]
  Now $u$ is the subsequence of $u^\top$ going from position $1$ to position $|u|$.  
  So Lemma \ref{LemAdequacy} tells us that we must have
  \[
    \blank,()\ts M \converges_u \skipp,()
    \]
  in \IAX.
\end{proof}

@@

\begin{theorem}[Full abstraction]
  Let $M,N\from T$ be closed terms of \IAX.  
  Then $M,N$ are observationally equivalent -- i.e., for all contexts $C[-]\from\com$ of \IAX with a hole of type $T$ and for all sequences $u\in X^*$, 
  \[
    \blank,()\ts C[M] \converges_u \skipp,() \Longleftrightarrow \blank,()\ts C[N] \converges_u \skipp,()\,\text{--}
    \]
  if and only if $\deno{M}\sim\deno{N}$.
  \label{TheKleisliFullAbstraction}
\end{theorem}
\begin{proof}
  First, suppose that $\deno{M}\sim\deno{N}$.  
  Let $C[-]\from\com$ be a context with a hole of type $T$.  
  Then the denotation of $t\ts C[t]$ is a morphism $\alpha\from \deno{T} \to \bC$.
  Moreover, the denotation of $C[M]$ is the composite $\deno{M};\alpha$ and that of $C[N]$ is the composite $\deno{N};\alpha$ by functional completeness.

@@

\begin{corollary}
  Let $M\from\com$ be a closed term of \IAX.  
  Consider the denotation $\deno M\from 1 \to \bC$ in $\G_X$ as a morphism $1 \to (X \to \bC)$ in $\G$.  
  Then there exists some sequence $u\in X^*$ such that the composite
  \[
    1 \xrightarrow{\deno M}
    (X \to \bC) \xrightarrow{\eta_u}
    (\Var \to \bN) \xrightarrow{\deno{\neww}}
    \bN \xrightarrow{t_{|u|}}
    \bC
    \]
  is not equal to $\bot$, if and only if
  \[
    \blank,()\ts M \converges \skipp,()\,.
    \]
  \label{CorMayAdequacy}
\end{corollary}

@@

\begin{theorem}
  Let $M,N\from T$ be closed terms of \IAX. Then $M,N$ are may-observationally equivalent -- i.e., for all contexts $C[-]\from \com$ of \IAX with a hole of type $T$, \[
    \blank,()\ts C[M] \converges \skipp,() \Leftrightarrow \blank,()\ts C[N] \converges \skipp,()\,\text{--}
    \]
  if and only if $\deno M \sim_{\may} \deno N$.
\end{theorem}
\begin{proof}
  Suppose that $\deno M\sim_{\may}\deno N$.  
  Let $C[-]$ be a context of \IAX, interpreted as a morphism $\alpha \from \deno T \to \bC$.

  Suppose that $\blank,()\ts C[M] \converges \skipp,()$.  
  So there is some sequence $u\in X^*$ such that $\blank,()\ts C[M] \converges_u \skipp,()$ and therefore the composite
  \[
    1 \xrightarrow{\deno M;\alpha}
    (X \to \bC) \xrightarrow{\eta_u}
    (\Var \to \bN) \xrightarrow{\deno{\neww}}
    \bN \xrightarrow{t_{|u|}}
    \bC
    \]
  is not equal to $\bot$.  
  Therefore, there exists some $v\in X^*$ such that the composite
  \[
    1 \xrightarrow{\deno N;\alpha}
    (X \to \bC) \xrightarrow{\eta_v}
    (\Var \to \bN) \xrightarrow{\deno{\neww}}
    \bN \xrightarrow{t_{|v|}}
    \bC
    \]
  Therefore, $\blank,()\ts C[N] \converges_v \skipp,()$ and so $\blank,()\ts C[N] \converges \skipp,()$.
  The reverse direction is identical.

@@

\begin{example}
  $\blank,()\ts \Y(\lambda x.x)\diverges$ because it has the following infinite derivation tree.
  \[
    \inferrule*
    {
      \inferrule*
      {
        \inferrule*
        {
        }
        {
          \blank,()\ts (\lambda x.x) \converges (\lambda x.x),()
        }
        \\
        \inferrule*
        {
          \vdots
        }
        {
          \blank,() \ts \Y(\lambda x.x)\diverges
        }
      }
      {
        \blank,()\ts (\lambda x.x)(\Y(\lambda x.x)) \diverges
      }
    }
    {
      \blank,()\ts \Y(\lambda x.x)\diverges
    }
    \]
\end{example}

@@

\begin{corollary}
  Let $M\from \com$ be a closed term of \IAX and consider the denotation $\deno M \from 1 \to \bC$ in $\G_X$ as a morphism $1 \to (X \to \bC)$ in $\G$.
  Then $\blank,()\ts M\mustconverge$ if and only if $\deno{M}\downarrow_{\must}$.
  \label{CorAdeqMust}
\end{corollary}

@@

We have shown the following adequacy result.
\begin{corollary}
  Let $M\from \com$ be a closed term of \IAX and consider the denotation $\deno M \from 1 \to \bC$ in $\G_X$ as a morphism $X \to \bC$ in $\G$.
  Then $\blank,()\ts M\mustconverge$ if and only if this strategy is winning.
\end{corollary}
